Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics

### 2024-08-25 Thesis/03-chap3.qmd
modify reason for using scagnostics

Providing multiple residual plots to the model, such as a pair of plots, a triplet or a lineup is also a possible option. @chopra2005learning have shown that computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar. Applied to our specific problem, we can define null plots of a fitted regression model to be similar to each other, while considering true residual plots to be distinct from null plots of any fitted regression model. A triplet constitutes a set of three images, denoted as $image_1$, $image_2$ and $image_3$. It is often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings b
etween samples. For this setup, we can apply the same criteria to define similarity between images. However, it is important to note that these two approaches usually require additional considerations regarding the loss function and, at times, non-standard training processes due to shared weights between different convolutional blocks.

@@

>>>>

Providing multiple residual plots to the model, such as a pair of plots, a triplet or a lineup is also a possible option. @chopra2005learning have shown that computer vision models designed for image comparison can assess whether a pair of images are similar or dissimilar. Applied to our specific problem, we can define null plots of a fitted regression model to be similar to each other, while considering true residual plots to be distinct from null plots of any fitted regression model. A triplet constitutes a set of three images, denoted as $image_1$, $image_2$ and $image_3$. It is often used to predict whether $image_2$ or $image_3$ is more similar to $image_1$, proving particularly useful for establishing rankings between samples. For this setup, we can apply the same criteria to define similarity between images. However, it is important to note that these two approaches usually require additional considerations regarding the loss function and, at times, non-standard training processes due to shared weights between different convolutional blocks.

@@

For each residual plot used as an input image, we calculated four scagnostics — "Monotonic", "Sparse", "Splines", and "Striped" – using the `cassowaryr` R package [@mason2022cassowaryr]. These computed measures, along with the number of observations from the fitted model, were provided as the second input for the computer vision model. We selected these scagnostics due to their reliability and efficiency, as other scagnostics occasionally caused R process crashes (approximately 5% of the time) during training data preparation, due to a bug in the `interp` R package [@Albrecht2023interp]. Although the package maintainer later fixed this bug at our request, the fix came too late to retrain the model, and additionally, their high computational costs make them unsuitable for rapid inference, which was a critical factor in our choice.
